ICBEdataset figure Metro Maps

### 2022-07-18 replication_paper/pnas_draft/paper.Rmd
v1.1 first self contained hopefully

icews_clean_471 <- read_tsv("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/data/out/icews_clean_471.tsv")
dim(icews_clean_471) #270,642
icews_clean_471_small <- read_tsv("/mnt/8tb_a/rwd_github_private/ICBEdataset/replication_paper/data/out/icews_clean_471_small.tsv")

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ ICEWs Counts of Events per Day by Type during the Crimea-Donbas Crisis \label{fig:p_precision_icews}}
\centering{\includegraphics[width=17cm]{../figures/p_precision_icews.png}}
\textit{Notes:Unit of analysis is the Dyad-Day. Edges <-> indicates undirected dyad and -> indicates directed dyad. Top 10 most active dyads per category shown. Red text shows events from the synthetic narrative relative to that event category. Event types are grouped by hand along similar themes. Blue bars indicate an event recorded by ICEWs for that dyad on that day. }
\end{figure}
\clearpage
\twocolumn

>>>>

@@

\clearpage
\onecolumn
\begin{figure}[H]
\caption{ ICEWs Events by Day by Type during the Crimea-Donbas Crisis \label{fig:p_precision_icews}}
\centering{\includegraphics[width=17cm]{../figures/p_precision_icews.png}}
\textit{Notes:Unit of analysis is the Dyad-Day. Edges <-> indicates undirected dyad and -> indicates directed dyad. Top 10 most active dyads per category shown. Red text shows events from the synthetic narrative relative to that event category. Blue bars indicate an event recorded by ICEWs for that dyad on that day. }
\end{figure}
\clearpage
\twocolumn

### 2022-07-18 replication_paper/pnas_draft/paper.tex
v1.1 first self contained hopefully

Each crisis was typically assigned to 2 expert coders and 2 novice coders with an additional tie-breaking expert coder assigned to sentences with high disagreement.\footnote{Expert coders were graduate students or postgraduates who collaboratively developed the ontology and documentation for the codebook. Undergraduate coders were students who engaged in classroom workshops.} For the purposes of measuring intercoder agreement and consensus, we temporarily disaggregate the unit of analysis to the Coder-Crisis-Sentence-Tag (n=971,547), where a tag is any unique piece of information a coder can associate with a sentence such as an actor, date, behavior, etc. We then aggregate those tags into final events (n=18,309), using a consensus procedure (SI Appendix, Algorithm B2) that requires a tag to have been chosen by at least one expert coder and either a majority of expert or novice coders. This screens noisy tags that no expert considered possible but leverages novice knowledge to tie-break between equally plausible tags chosen by experts.

@@

We evaluate the internal validity of the coding process in several ways. For every tag applied we calculate the observed intercoder agreement as the percent of other coders who also applied that same tag (SI Appendix, Fig. B3). Across all concepts, the Top 1 Tag Agreement was low among novices (34\%), moderate for experts (65\%), and high (87\%) following the consensus screening procedure.

@@

The other side of event measurement is precision, the degree to which a sequence of events correctly and usefully describes the information in history, \(Pr(H|E)\). It does little good to recall a historical event but too vaguely (e.g.~MIDs describes the Cuban Missile crisis as a blockade, a show of force, and a stalemate) or with too much error (e.g.~ICEWS records 263 ``Detonate Nuclear Weapons'' events between 1995-2019) to be useful for downstream applications. ICBe's ontology and coding system is designed to strike a balance so that the most important information is recovered accurately but also abstracted to a level that is still useful and interpretable. You should be able to lay out events of a crisis on a timeline, as in Figure 3, and read off the macro structure of an episode from each individual move. We call this visualization a crisis map, a directed graph interesected with a timeline, and provide crisis maps for every event dataset for each case study (SI Appendix, Fig. D3 and D4) and all crises on the companion website.

@@

We take the same approach here in Figure 5, selecting four specific CAMEO event codings and checking how often they reflect a true real world event. We choose four event types around key moments in the crisis. The start of the crisis revolves around the Ukraine backing out of trade deal with the EU in favor of Russia, but the ``sign formal agreement'' acts more like a topic detector with dozens of events generated by discussions of a possible agreement but not the actual agreement which never materialized. The switch is caught by the ``reject plan, agreement to settle dispute'', but also continues for Victor Yanukovych for even after he was removed from power because of articles retroactively discussing the cause of his of his removal. Events for ``use conventional military force'' capture a threshold around the start of hostilities and who the participants were but not any particular battles or campaigns. Likewise, ``impose embargo, boycott, or sanctions'' captures the start of waves of sanctions and from who but are effectively costantly as the news coverage does not distinguish between subtle changes or additions. In sum, dictionary based methods on news corpora tend to have high recall because they parse everything in the news, but for the same reason their specificity for most event types is too low to back out individual chess like sequencing that ICBe aims to record.

>>>>

Each crisis was typically assigned to 2 expert coders and 2 novice coders with an additional tie-breaking expert coder assigned to sentences with high disagreement.\footnote{Expert coders were graduate students or postgraduates who collaboratively developed the ontology and documentation for the codebook. Undergraduate coders were students who engaged in classroom workshops.} For the purposes of measuring intercoder agreement and consensus, we temporarily disaggregate the unit of analysis to the Coder-Crisis-Sentence-Tag (n=972,545), where a tag is any unique piece of information a coder can associate with a sentence such as an actor, date, behavior, etc. We then aggregate those tags into final events (n=18,309), using a consensus procedure (SI Appendix, Algorithm B2) that requires a tag to have been chosen by at least one expert coder and either a majority of expert or novice coders. This screens noisy tags that no expert considered possible but leverages novice knowledge to tie-break between equally plausible tags chosen by experts.

@@

We evaluate the internal validity of the coding process in several ways. For every tag applied we calculate the observed intercoder agreement as the percent of other coders who also applied that same tag (SI Appendix, Fig. B3). Across all concepts, the Top 1 Tag Agreement was low among novices (31\%), moderate for experts (65\%), and high (74\%) following the consensus screening procedure.

@@

The other side of event measurement is precision, the degree to which a sequence of events correctly and usefully describes the information in history, \(Pr(H|E)\). It does little good to recall a historical event but too vaguely (e.g.~MIDs describes the Cuban Missile crisis as a blockade, a show of force, and a stalemate) or with too much error (e.g.~ICEWS records 263 ``Detonate Nuclear Weapons'' events between 1995-2019) to be useful for downstream applications. ICBe's ontology and coding system is designed to strike a balance so that the most important information is recovered accurately but also abstracted to a level that is still useful and interpretable. You should be able to lay out events of a crisis on a timeline, as in Figure 3, and read off the macro structure of an episode from each individual move. We call this visualization a crisis map, a directed graph intersected with a timeline, and provide crisis maps for every event dataset for each case study (SI Appendix, Fig. D3 and D4) and all crises on the companion website.

@@

We take the same approach here in Figure 5, selecting four specific CAMEO event codings and checking how often they reflect a true real world event. We choose four event types around key moments in the crisis. The start of the crisis revolves around Ukraine backing out of trade deal with the EU in favor of Russia, but ``sign formal agreement'' events act more like a topic detector with dozens of events generated by discussions of a possible agreement but not the actual agreement which never materialized. The switch is caught by the ``reject plan, agreement to settle dispute'', but also continues for Victor Yanukovych for even after he was removed from power because of articles retroactively discussing the cause of his of his removal. Events for ``use conventional military force'' capture a threshold around the start of hostilities and who the participants were but not any particular battles or campaigns. Likewise, ``impose embargo, boycott, or sanctions'' captures the start of waves of sanctions and from who but are effectively constantly as the news coverage does not distinguish between subtle changes or additions. In sum, dictionary based methods on news corpora tend to have high recall because they parse everything in the news, but for the same reason their specificity for most event types is too low to back out individual chess like sequencing that ICBe aims to record.

### 2022-07-18 replication_paper/rmd/old/06_ICBEdataset_figure_metro_maps.Rmd
v1.1 first self contained hopefully

@@

@@

```{r}
#Maybe it's just ISO3 country codes?
terrier_clean_temp <- terrier %>% 
    dplyr::select(
      cameo_code=x3,
      src_country=x8,
      src_agent=x9,
      src_other_agent=x10,

      year=x17,
      month=x18,
      day=x19#,
      #random_text=x28
    ) %>% 
    filter(!is.na(src_country) & !is.na(target_country)) %>%
    distinct() %>% 
     mutate(cameo_code=cameo_code %>% as.numeric()) %>% 
     left_join(CAMEO_eventcodes %>% mutate(cameo_code=cameo_code %>% as.numeric()) %>%
                 dplyr::select(cameo_code=cameo_code, cameo_label= cameo_label)  ) %>%
    mutate( cameo_label= cameo_label %>% str_replace(', not specified below','')) %>%
    distinct()

#done

```{r}
terrier_clean_temp <- terrier_clean_temp %>%

@@

   left_join(df %>% dplyr::select(src_other_agent=agent   , src_other_agent_label=agent_label) %>% distinct() %>% na.omit() ) %>%
   left_join(df %>% dplyr::select(target_other_agent=agent, target_other_agent_label=agent_label) %>% distinct() %>% na.omit() )

@@

```{r}
terrier_clean <- terrier_clean_temp %>%
     left_join(df %>% dplyr::select(src_agent=agent, src_agent_label=agent_label) %>% distinct() %>% na.omit() ) %>%
   left_join(df %>% dplyr::select(target_agent=agent, target_agent_label=agent_label) %>% distinct() %>% na.omit() )  %>% 
   distinct() %>%
   filter(
     !src_other_agent_label %in% role_rejects &
     !target_other_agent_label %in% role_rejects &
     !src_agent_label %in% role_rejects &
     !target_agent_label %in% role_rejects 
     ) %>% 
  left_join(actor_translator %>% dplyr::select(src_country=terrier,actor_1_qcode=QCode,actor_1=QCode_name) ) %>%
  left_join(actor_translator %>% dplyr::select(target_country=terrier,actor_2_qcode=QCode,actor_2=QCode_name) ) %>%
  filter(!is.na(actor_1_qcode) & !is.na(actor_2_qcode) ) %>%
  mutate(source_root_target=paste0(src_country, "_", target_country) ) %>%
  mutate(
  event_date= lubridate::as_date(paste0( year, "-", month,"-",day ))
  )
dim(terrier_clean) #2,049,759

#For now
#IGO 223751 for now we're going to exclue just generic Intergovernmental organization and multiinational corps
#MNC 62721
#terrier_clean %>% dplyr::select(src_country, actor_1) %>% filter(is.na(actor_1)) %>% count(src_country) %>% View()

@@

@@

terrier_metro_plots <- function(crisis=196, topactor=F){

@@

@@

                           filter(
                                  source_root_target %in% diads_qcodes$source_root_target & 
                                  #target_root %in% na.omit(actors_translator_subset$phoenix_2020) & 
                                  #source_root %in% na.omit(actors_translator_subset$phoenix_2020) &
                                  event_date>=date_start &
                                  event_date<=date_end   
                                  #date_simple>=start_simple & 
                                  #date_simple<=end_simple 
                                  ) %>%

>>>>

if(file.exists(paste0(here::here(), "/replication_paper/data/ignore/terrier/largegeolocateddata//"))){
  terrier_fromscratch<-readline(prompt="Load original Terrier data? (T/F)")
}

if(terrier_fromscratch){

@@

}
```

if(terrier_fromscratch){
#Maybe it's just ISO3 country codes?
terrier_clean_temp <- terrier %>% 
    dplyr::select(
      cameo_code=x3,
      src_country=x8,
      src_agent=x9,
      src_other_agent=x10,

      year=x17,
      month=x18,
      day=x19#,
      #random_text=x28
    ) %>% 
    filter(!is.na(src_country) & !is.na(target_country)) %>%
    distinct() %>% 
     mutate(cameo_code=cameo_code %>% as.numeric()) %>% 
     left_join(CAMEO_eventcodes %>% mutate(cameo_code=cameo_code %>% as.numeric()) %>%
                 dplyr::select(cameo_code=cameo_code, cameo_label= cameo_label)  ) %>%
    mutate( cameo_label= cameo_label %>% str_replace(', not specified below','')) %>%
    distinct() 
}
#done

@@

```{r}
if(terrier_fromscratch){
terrier_clean_temp <- terrier_clean_temp %>%

@@

   left_join(df %>% dplyr::select(src_other_agent=agent   , src_other_agent_label=agent_label) %>% distinct() %>% na.omit() ) %>%
   left_join(df %>% dplyr::select(target_other_agent=agent, target_other_agent_label=agent_label) %>% distinct() %>% na.omit() ) 
}

@@

```{r}
if(terrier_fromscratch){
terrier_clean <- terrier_clean_temp %>%
     left_join(df %>% dplyr::select(src_agent=agent, src_agent_label=agent_label) %>% distinct() %>% na.omit() ) %>%
   left_join(df %>% dplyr::select(target_agent=agent, target_agent_label=agent_label) %>% distinct() %>% na.omit() )  %>% 
   distinct() %>%
   filter(
     !src_other_agent_label %in% role_rejects &
     !target_other_agent_label %in% role_rejects &
     !src_agent_label %in% role_rejects &
     !target_agent_label %in% role_rejects 
     ) %>% 
  left_join(actor_translator %>% dplyr::select(src_country=terrier,actor_1_qcode=QCode,actor_1=QCode_name) ) %>%
  left_join(actor_translator %>% dplyr::select(target_country=terrier,actor_2_qcode=QCode,actor_2=QCode_name) ) %>%
  filter(!is.na(actor_1_qcode) & !is.na(actor_2_qcode) ) %>%
  mutate(source_root_target=paste0(src_country, "_", target_country) ) %>%
  mutate(
  event_date= lubridate::as_date(paste0( year, "-", month,"-",day ))
  )
dim(terrier_clean) #2,049,759

saveRDS(terrier_clean, here::here("replication_paper", "data", "in", "terrier_clean.Rds"))
}

#For now
#IGO 223751 for now we're going to exclude just generic Intergovernmental organization and multinational corps
#MNC 62721
#terrier_clean %>% dplyr::select(src_country, actor_1) %>% filter(is.na(actor_1)) %>% count(src_country) %>% View()

@@

if(!terrier_fromscratch){
  terrier_clean <- readRDS(here::here("replication_paper", "data", "in", "terrier_clean.Rds"))
}

@@

terrier_metro_plots <- function(crisis=426, topactor=F){

@@

  date_start <- lubridate::as_date(paste0(icb_crises[icb_crises$crisno==crisis,]$yrtrig , "-", icb_crises[icb_crises$crisno==crisis,]$motrig, "-01"))
  date_end <- lubridate::as_date(paste0(icb_crises[icb_crises$crisno==crisis,]$yrterm , "-", icb_crises[icb_crises$crisno==crisis,]$moterm+1, "-01"))

@@

                           dplyr::filter(
                                  source_root_target %in% diads_qcodes$source_root_target & 
                                  #target_root %in% na.omit(actors_translator_subset$phoenix_2020) & 
                                  #source_root %in% na.omit(actors_translator_subset$phoenix_2020) &
                                  event_date>=date_start &
                                  event_date<=date_end   
                                  #date_simple>=start_simple & 
                                  #date_simple<=end_simple 
                                  ) %>%
